```{r}
fortune_1000 <- read.csv('/Users/sandeep/Downloads/DSRP_2023_AGD/csv_files/Fortune_1000.csv')
big_fortune_1000 <- mutate(fortune_1000, prof_margin = profit/revenue * 100, reduced_revenue = revenue/1000)
companies_in_MN <- select(filter(big_fortune_1000, state == "MN"), company, prof_margin, state)
companies_in_US <- select(filter(big_fortune_1000, state != "MN"), company, prof_margin, state)
fortune_1000
```
## T-test Hypotheses

### Null Hypothesis: The profit margins of companies headquartered in Minnesota is equal to the profit margins of companies headquartered elsewhere in the U.S.

### Alternate Hypothesis: The profit margins of companies headquartered in Minnesota is less than the profit margins of companies headquartered elsewhere in the U.S.

```{r}
# t.test(col1, col2, paired = T/F, alternative, conf_lvl, mu)

t_test_val <- t.test(companies_in_MN$prof_margin, na.omit(companies_in_US$prof_margin))
t_test_val
```
## ANOVA in R using 'aov'

```{r}
#aov(data = df, numeric_var ~ categorical_var)

anova_results <- aov(data = iris, Sepal.Length ~ Species)
summary(anova_results)
TukeyHSD(anova_results)
```
## ANOVA Hypotheses

### Null Hypothesis: The average number of employees in every sector is equal

### Alternative Hypothesis: The average number of employees in every sector is not equal

```{r}
aov_results <- aov(data = fortune_1000, num..of.employees ~ sector)
summary(aov_results)
```

## Chi Square Hypotheses

### Null Hypothesis: There is no relationship between whether a company is profitable and the state it is headquartered in.

### Alternative Hypothesis: There is a relationship between whether a company is profitable and the state it is headquartered in.

```{r}
library(corrplot)
t <- table(fortune_1000$state, fortune_1000$profitable)
t

chi <- chisq.test(t)
chi$p.value
chi$residuals

corrplot(chi$residuals, is.cor = F)
```

